---
title: "Clustering & processing report"
date: "`r format(Sys.time(), '%Y/%m/%d %H:%M:%S')`"
output: html_document
editor_options:
  chunk_output_type: inline
params:
  wd: NULL
  config: NULL
  assets: NULL
  prefix: NULL
  experiment_id: NULL
---

```{r load_libraries, cache = FALSE, message = FALSE, warning = FALSE}

# libraries from CRAN
library(readr)
library(ggplot2)
library(Seurat)
library(clustree)
library(purrr)
library(SCpubr)
```

Read in the config file for this run. These parameters should be located in ~/assets and contain all information needed for the run

```{r read_config, message = FALSE, warning = FALSE}
config <- read_delim(file.path(params$assets,'clustering.config.tsv'), delim = '\t', escape_double = FALSE, trim_ws = TRUE)

# coerce to a list of param:value pairs, and make sure numeric parameters are numeric type
config <- as.list(tibble::deframe(config[, c(1, 2)]))

# automatically convert each config element to the right type
config <- lapply(config, type.convert, as.is = TRUE)

config

info_samples <- read_tsv(file.path(params$assets,"integration.seurat.tsv"))

# convert to a list of colour palettes
info_groups <- read_tsv(file.path(params$assets,"integration.covariates.tsv")) %>% 
  split(f = .$covariate) %>%
  map(~ .x %>% 
        dplyr::select(-covariate) %>% 
        tibble::deframe()
      )
# make the sample palette a covariate
info_groups$RunID <- info_samples %>% 
  dplyr::select(sample, color) %>% 
  tibble::deframe()
```

Set up directory structure and confirm current working directory:

```{r setup_dir}
# print current working directory
print(getwd())

# specify and create output and figures folders with run prefix
output_dir <- paste0(params$prefix,'_output')
if (dir.exists(file.path(params$wd,output_dir))) {
  cat("output dir already exists, overwriting data")
} else {
  cat("output dir does not yet exists, creating dir at\n", file.path(params$wd,output_dir))
  dir.create(file.path(params$wd,output_dir), showWarnings = FALSE, recursive = TRUE)
}

figures_dir <- paste0(params$prefix,'_figures')
if (dir.exists(file.path(params$wd,figures_dir))) {
  cat("figures dir already exists, overwriting data")
} else {
  cat("figures dir does not yet exists, creating dir at\n", file.path(params$wd,figures_dir))
  dir.create(file.path(params$wd,figures_dir), showWarnings = FALSE, recursive = TRUE)
}
```

Set up R markdown:

```{r setup}

knitr::opts_chunk$set(
  # Display PNG in HTML file, but also keep PDF outputs in figures folder
  dev = c("png", "pdf"), 
  # Keep all the figures produced in a chunk
  fig.keep = "all",      
  # Save all figures to this output folder
  fig.path = file.path(params$wd, paste0(params$prefix,'_figures/')),
  # Do not cache results
  cache = FALSE)

# not used
# cache.path = file.path(knitr::opts_knit$get("root.dir"), ".cache/")

# don't use dingbats in PDFs, in order to create Illustrator-friendly figures
grDevices::pdf.options(useDingbats = FALSE)

```
# Initialization and initial filtering

## Load Seurat object

```{r read seurat object}

seurat <- readRDS(config$rds_dir)
seurat
```

## Identify neighbours and cluster in different resolutions
```{r SNN}
# step 1. identify neighbours
seurat <- FindNeighbors(seurat,
                        reduction   = "runid.scvi",
                        dims        = 1:10, 
                        compute.SNN = TRUE,
                        verbose     = TRUE)
```

All parameters metrics are saved within the seurat object in the `seurat@misc` slot.
```{r clustering_resolution, warning=FALSE}

#step 2. find clusters in multiple resolutions
lowest.res <- 0
highest.res <- 3
stepsize <- 0.2

resolution.range <- seq(from = lowest.res, to = highest.res, by = stepsize) 
algorithm <- "louvain"

seurat@misc$parameters$clustering$resolution.range = resolution.range
seurat@misc$parameters$clustering$algorithm = algorithm

for (i in 1:length(resolution.range)){ 
  print(resolution.range[i])
  seurat <- FindClusters(seurat, 
                         resolution = resolution.range[i], 
                         algorithm = 1, 
                         verbose = F, 
                         random.seed = config$seed)
}
```

## Plot all clusters in UMAP space and as a clustertree

```{r runid_clustering_resolution_plots, fig.height=10, fig.width=10}
p5 <- list()
for (i in 1:length(resolution.range)) {
    p <- do_DimPlot(seurat,
                    reduction = 'runid.umap', 
                    group.by = paste0('RNA_snn_res.',resolution.range[i]), 
                    label = T, 
                    repel = T,
                    plot.title = paste0('RNA_snn_res.',resolution.range[i]), 
                    pt.size = 0.2, 
                    legend.position = "none") + coord_fixed()
    
    p5[[i]] <- p
}

walk(p5, print)
```


```{r clustree, fig.width=15, fig.height=20}
clustree(seurat)
```

# Save seurat 

```{r save seurat}
saveRDS(seurat, file = file.path(output_dir,paste0(params$prefix,"_seurat.rds")))
```

# Reproducibility

This document was last rendered on:

```{r time, echo = FALSE, cache = FALSE}

message(Sys.time())

```

The R session info:

<details>

```{r sinfo, cache = FALSE, echo = FALSE}

sessionInfo()

```

</details>
  