---
title: "`r params$title`"
author: "riri"
date: "2023-09-20"
output: html_document
editor_options: 
  chunk_output_type: inline
params:
  title: NULL
  wd: NULL
  assets: NULL
  prefix: NULL
---

```{r}
library(Seurat)
library(tidyverse)
library(readr)
```

Set up directory structure and confirm current working directory:

```{r setup_dir}
# print current working directory
getwd()

# specify and create output and figures folders with run prefix
output_dir <- file.path(params$wd, paste0(params$prefix,'_output/'))
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

figures_dir <- file.path(params$wd,paste0(params$prefix,'_figures/'))
dir.create(figures_dir, showWarnings = TRUE, recursive = TRUE)
```

Set up R markdown:

```{r setup}

knitr::opts_chunk$set(
  # Display PNG in HTML file, but also keep PDF outputs in figures folder
  dev = c("png", "pdf"), 
  # Keep all the figures produced in a chunk
  fig.keep = "all",      
  # Save all figures to this output folder
  fig.path = figures_dir,
  # Do not cache results
  cache = FALSE)

# not used
# cache.path = file.path(knitr::opts_knit$get("root.dir"), ".cache/")

# don't use dingbats in PDFs, in order to create Illustrator-friendly figures
grDevices::pdf.options(useDingbats = FALSE)

```

```{r read in config file}
config <- read_delim(file.path(params$assets,'demux.config.tsv'), delim = '\t', escape_double = FALSE, trim_ws = TRUE)

# coerce to a list of param:value pairs, and make sure numeric parameters are numeric type
config <- as.list(tibble::deframe(config[, c(1, 2)]))

# automatically convert each config element to the right type
config <- lapply(config, type.convert, as.is = TRUE)

config
```

```{r read 10X data}
matrices <- Read10X(config$tenX_dir)

umis <- matrices$`Gene Expression`
htos <- matrices$`Antibody Capture`
```

filter cells that are detected in both RNA and HTO
```{r filter cells}
joint.bcs <- intersect(colnames(umis), colnames(htos))

umis <- umis[,joint.bcs]
htos <- as.matrix(htos[,joint.bcs])

rownames(htos)
```

Create the seurat object and add HTO data
```{r}
seurat <- CreateSeuratObject(counts = umis)

seurat[["HTO"]] <- CreateAssayObject(counts = htos)
seurat <- NormalizeData(seurat, assay = "HTO", normalization.method = "CLR")
```

Demultiplex cells and visualize the results

```{r demux}
seurat <- HTODemux(seurat, assay = "HTO", positive.quantile = config$pos.quantile, seed = config$seed_use)
table(seurat$HTO_classification.global)
table(seurat$HTO_classification)
```
```{r HTO_Enrichments_1, fig.height=15, fig.width=10}
Idents(seurat) <- "HTO_maxID"
RidgePlot(seurat, assay = "HTO", features = rownames(seurat[["HTO"]]))
```

```{r HTO_Enrichments_2, fig.height=5, fig.width=10}
Idents(seurat) <- "HTO_maxID"
HTOHeatmap(seurat, assay = "HTO")
Idents(seurat) <- "HTO_classification.global"
VlnPlot(seurat, features = "nCount_RNA", log = T)
```

subset the seurat on singlets and check for batch effects

```{r Batch_effects, fig.height=5, fig.width=10}

# Extract the singlets
singlets <- subset(seurat, idents = "Singlet")

singlets <- NormalizeData(singlets)
# Select the top 1000 most variable features
singlets <- FindVariableFeatures(singlets, selection.method = "mean.var.plot")

# Scaling RNA data, we only scale the variable features here for efficiency
singlets <- ScaleData(singlets, features = VariableFeatures(singlets))

# Run PCA
singlets <- RunPCA(singlets, features = VariableFeatures(singlets))

# We select the top 10 PCs for clustering and tSNE based on PCElbowPlot
singlets <- RunTSNE(singlets, reduction = "pca", dims = 1:20)

# Projecting singlet identities on TSNE visualization
SCpubr::do_DimPlot(singlets, group.by = "HTO_classification", reduction = "tsne") + coord_fixed()
```


Saves non-processed seurat object with singlets from HTODemux
Input for preprocessing pipeline
```{r cleaning & saving}

seurat <- subset(seurat, idents = "Singlet")
seurat$sample <- seurat$HTO_classification

HTO_meta <- seurat@meta.data

seurat@meta.data[,4:11] <- NULL

saveRDS(seurat, file = file.path(output_dir, paste0(params$title,"_seurat.rds")))
write_tsv(HTO_meta, file = file.path(output_dir, paste0(params$title,"_HTO_meta.tsv")))
```

```{r}
sessionInfo()
```